Comparing Power Optimizations in the Reorder Buffer

### 2011-06-07 paper/paper.tex
Buggy.  Henry helping.
Finished the instruction generation section.

\section{Introduction}
BLAH BLAH BLAH INTRO WEEEE
\section{Project Implementation}
	Our simulator provides a cycle accurate behavioral model of a standard circular re-order buffer.  At the outset of the project, the standard re-order buffer we intended to implement was a non-circular RAM-based re-order buffer.  However, after feedback from the course instructor and students it was decided that a circular buffer is more widely held as being the 'standard' buffer type and would serve as an ideal underlying model for power optimizations.  Thus, we implement and compare the standard circular re-order buffer to circular re-order buffers with the following optimizations: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch). The following sections outline each of the above model implementations as well as input instruction generation and power results tabulation.  
\subsection{Instruction Generation}
	We model the input to our simulator as a stream of integer and floatpoint instructions.  The instructions are represented at the bit-level and model MIPS instructions as closely as possible.  Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.  Table XX indicates the types of instructions modeled.  Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.  As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.  The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator[REFXX], with the exception of LOAD and STORE.  The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.  This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.  Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.
\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|} \hline
Instruction Type&Cycle Count&Instruction Type&Cycle Count\\ \hline
\texttt{LOAD}&20&\texttt{STORE}&20\\ \hline
\texttt{ADD}&1&\texttt{FADD}&3\\ \hline
\texttt{SUB}&1&\texttt{FSUB}&3\\ \hline
\texttt{MULT}&3&\texttt{FMULT}&5\\ \hline
\texttt{DIV}&18&\texttt{FDIV}&6\\ \hline\end{tabular}
\caption{Modeled instruction types and their corresponding completion cycle counts.}
\end{table*}

@@

	For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.  Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.  The specific values used for each workload are listed in Table XX.

\subsection{Model Implementation}
\subsubsection{Circular ROB}
\subsubsection{Distributed ROB}
\subsubsection{Dynamic ROB}
\subsubsection{Reduced Output Port ROB (Retention Latch)}
\subsection{Power Tabulation Implementation}
\section{Results}
hi
\subsection{subsection titl}
sub section.. with table!
\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

>>>>

\section{Introduction}
BLAH BLAH BLAH INTRO WEEEE
\section{Project Implementation}
Our simulator provides a cycle accurate behavioral model of a standard circular re-order buffer.  At the outset of the project, the standard re-order buffer we intended to implement was a non-circular RAM-based re-order buffer.  However, after feedback from the course instructor and students it was decided that a circular buffer is more widely held as being the 'standard' buffer type and would serve as an ideal underlying model for power optimizations.  Thus, we implement and compare the standard circular re-order buffer to circular re-order buffers with the following optimizations: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch). The following sections outline each of the above model implementations as well as input instruction generation and power results tabulation.  
\subsection{Instruction Generation}

We model the input to our simulator as a stream of integer and floatpoint instructions.  The instructions are represented at the bit-level and model MIPS instructions as closely as possible.  Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.  Table XX indicates the types of instructions modeled.  Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.  As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.  The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator[REFXX], with the exception of LOAD and STORE.  The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.  This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.  Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.
\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|} \hline
Instruction Type&Cycle Count&Instruction Type&Cycle Count\\ \hline
\texttt{LOAD}&20&\texttt{STORE}&20\\ \hline
\texttt{ADD}&1&\texttt{FADD}&3\\ \hline
\texttt{SUB}&1&\texttt{FSUB}&3\\ \hline
\texttt{MULT}&3&\texttt{FMULT}&5\\ \hline
\texttt{DIV}&18&\texttt{FDIV}&6\\ \hline\end{tabular}
\caption{Modeled instruction types and their corresponding completion cycle counts.}
\end{table*}

@@

For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.  Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.  The specific values used for each workload are listed in Table XX.
\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|c|} \hline
ins{\char'137}profile&default{\char'137}profile&173{\char'056}applu&191{\char'056}fma3d&183{\char'056}equake\\ \hline
loadProc&0.10&0.32&0.12&0.37\\ \hline
storeProc&0.10&0.12&0.24&0.06\\ \hline
addProc&0.10&0.01&0.15&0.08\\ \hline
subProc&0.10&0.01&0.15&0.08\\ \hline
multProc&0.10&0.01&0.15&0.08\\ \hline
divProc&0.10&0.01&0.15&0.08\\ \hline
faddProc&0.10&0.13&0.01&0.07\\ \hline
fsubProc&0.10&0.13&0.01&0.06\\ \hline
fmultProc&0.10&0.13&0.01&0.06\\ \hline
fdivProc&0.10&0.13&0.01&0.06\\ \hline
forwardingRate&0.10&0.06&0.06&0.06\\ \hline
instCount&10000&10000&10000&10000\\ \hline\end{tabular}
\caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
\end{table*}
\subsection{Model Implementation}
\subsubsection{Circular ROB}
\subsubsection{Distributed ROB}
\subsubsection{Dynamic ROB}
\subsubsection{Reduced Output Port ROB (Retention Latch)}
\subsection{Power Tabulation Implementation}
\section{Results}
hi
\subsection{subsection titl}
sub section.. with table!
\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!
